---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo=TRUE}
unzip("activity.zip")
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```
  
  
  
  
## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
dfSum <- df %>% group_by(date) %>% 
      summarize(TotSteps = sum(steps, na.rm = TRUE)) 

ggplot(dfSum, aes(x = TotSteps)) +
  geom_histogram(bins = 20) +
  labs(x = "Total Steps", y = "Frequency", title = "Total Number of Steps Taken Each Day")

MeanTotSteps <- mean(dfSum$TotSteps)
MedianTotSteps <- median(dfSum$TotSteps)

```
The mean of the total number of steps taken per day: `r MeanTotSteps`  
The median of the total number of steps taken per day: `r MedianTotSteps`
  
  
  
  
## What is the average daily activity pattern?
```{r, echo=TRUE}
library(scales)

dfSum <- df %>% group_by(interval) %>% 
      summarize(MeanSteps = mean(steps, na.rm = TRUE))

dfSum$time <- sprintf("%04d", dfSum$interval)
dfSum$time <- as.POSIXct(strptime(dfSum$time, format = "%H%M"), tz="GMT")

ggplot(dfSum, aes(x = time, y= MeanSteps)) +
      geom_line() +
      labs(x = "Time Interval", y = "Average Steps", 
           title = "Number of Steps Taken, Averaged Across All Days") +
      scale_x_datetime(breaks = date_breaks("4 hour"), labels=date_format("%H:%M"))

MaxStepInt <- dfSum$interval[dfSum$MeanSteps == max(dfSum$MeanSteps)]
```
5-minute interval containing the maximum number of steps: `r MaxStepInt`   
  
  
  
  
## Imputing missing values
```{r, echo=TRUE}
TotNA <- sum(is.na(df$steps))

## Missing values filled with median values of corresponding 5-minutes interval
## Calculate median values for each interval
dfMed <- df %>% group_by(interval) %>% 
      summarize(MedSteps = median(steps, na.rm = TRUE))

## Add median values to data frame
dfImputed <- left_join(df, dfMed)

## find rows with missing values
na.id <- which(is.na(df$steps))

## Replace missing values by median values
dfImputed[na.id, "steps"] <- dfImputed[na.id, "MedSteps"]
dfImputed <- select(dfImputed, -MedSteps)

## Make a histogram and calculate mean and median values
dfSum <- dfImputed %>% group_by(date) %>% 
      summarize(TotSteps = sum(steps)) 

ggplot(dfSum, aes(x = TotSteps)) +
  geom_histogram(bins = 20) +
  labs(x = "Total Steps", y = "Frequency", title = "Total Number of Steps Taken Each Day")

MeanTotStepsImputed <- mean(dfSum$TotSteps)
MedianTotStepsImputed <- median(dfSum$TotSteps)

MeanImpact <- MeanTotStepsImputed - MeanTotSteps
MedianImpact <- MedianTotStepsImputed - MedianTotSteps
```
Total number of missing values in the dataset: `r TotNA`  
Impact of imputing missing data on the estimates of the total daily number of steps:  
- Mean Value: `r MeanImpact`  
- Median Value: `r MedianImpact`  
  
    
## Are there differences in activity patterns between weekdays and weekends?
```{r facet, fig.hight=4, echo=TRUE}
Sys.setlocale("LC_ALL","English")

dfImputed$Week <- factor(ifelse(weekdays(dfImputed$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))

dfSum1 <- dfImputed %>% group_by(Week, interval) %>% 
      summarize(MeanSteps = mean(steps))

dfSum1$time <- sprintf("%04d", dfSum1$interval)
dfSum1$time <- as.POSIXct(strptime(dfSum1$time, format = "%H%M"), tz="GMT")

ggplot(dfSum1, aes(x = time, y= MeanSteps)) +
      geom_line() +
      facet_grid(Week ~ .) +
      labs(x = "Time Interval", y = "Average Steps", 
           title = "Number of Steps Taken, Averaged Across All Days") +
      scale_x_datetime(breaks = date_breaks("4 hour"), labels=date_format("%H:%M"))
      
```
  
  